\subsection{Editor}

{\small
\begin{longtable}{|l|l|}\hline
\textbf{Function}		                    &   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Undo last action 	                        &   Ctrl+Z                  \\ \hline
Redo last action 	                        &   Ctrl+ Shift+Z           \\ \hline
Cut selected text                           &   Ctrl+X                  \\ \hline
Copy selected text                          &   Ctrl+C                  \\ \hline
Paste text from clipboard                   &   Ctrl+V                  \\ \hline
Select all text                             &   Ctrl+A                  \\ \hline
Swap header / source 	                    &	F11                     \\ \hline
Comment highlighted code                    &	Ctrl+Shift+C            \\ \hline
Uncomment highlighted code                  & 	Ctrl+Shift+X            \\ \hline
Duplicate line caret is on                  & 	Ctrl+D                  \\ \hline
Auto+complete / Abbreviations               & 	Ctrl+Space/Ctrl+J       \\ \hline
Show call tip 	                            &	Ctrl+Shift+Space        \\ \hline
Swap line caret is on with line above it    &	Ctrl+T                  \\ \hline
Toggle bookmark 	                        &	Ctrl+B                  \\ \hline
Goto previous bookmark 	                    &	Alt+PgUp                \\ \hline
Goto next bookmark 	                        &	Alt+PgDown              \\ \hline
Toggle current block folding 	            &	F12                     \\ \hline
Toggle all folds 	                        &	Shift+F12               \\ \hline
\caption{Basic shortcuts}
\end{longtable}
}

This is a list of shortcuts provided by the \codeblocks editor component. These shortcuts cannot be rebound.

{\small
\begin{longtable}{|l|l|}\hline
\textbf{Function}		                    &   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Magnify text size. 	                        &   Ctrl+Keypad "+"         \\ \hline
Reduce text size. 	                        &   Ctrl+Keypad "-"         \\ \hline
Restore text size to normal. 	            &   Ctrl+Keypad "/"         \\ \hline
Cycle through recent files. 	            &   Ctrl+Tab                \\ \hline
Indent block. 	                            &   Tab                     \\ \hline
Dedent block. 	                            &   Shift+Tab               \\ \hline
Delete to start of word. 	                &   Ctrl+BackSpace          \\ \hline
Delete to end of word. 	                    &   Ctrl+Delete             \\ \hline
Delete to start of line. 	                &   Ctrl+Shift+BackSpace    \\ \hline
Delete to end of line. 	                    &   Ctrl+Shift+Delete       \\ \hline
Go to start of document. 	                &   Ctrl+Home               \\ \hline
Extend selection to start of document. 	    &   Ctrl+Shift+Home         \\ \hline
Go to start of display line. 	            &   Alt+Home                \\ \hline
Extend selection to start of display line. 	&   Alt+Shift+Home          \\ \hline
Go to end of document. 	                    &   Ctrl+End                \\ \hline
Extend selection to end of document. 	    &   Ctrl+Shift+End          \\ \hline
Go to end of display line. 	                &   Alt+End                 \\ \hline
Extend selection to end of display line. 	&   Alt+Shift+End           \\ \hline
Expand or contract a fold point. 	        &   Ctrl+Keypad "*"         \\ \hline
Create or delete a bookmark. 	            &   Ctrl+F2                 \\ \hline
Go to next bookmark. 	                    &   F2                      \\ \hline
Select to next bookmark. 	                &   Alt+F2                  \\ \hline
Find selection. 	                        &   Ctrl+F3                 \\ \hline
Find selection backwards. 	                &   Ctrl+Shift+F3           \\ \hline
Scroll up. 	                                &   Ctrl+Up                 \\ \hline
Scroll down. 	                            &   Ctrl+Down               \\ \hline
Line cut. 	                                &   Ctrl+L                  \\ \hline
Line copy. 	                                &   Ctrl+Shift+T            \\ \hline
Line delete. 	                            &   Ctrl+Shift+L            \\ \hline
Line transpose with previous. 	            &   Ctrl+T                  \\ \hline
Line duplicate. 	                        &   Ctrl+D                  \\ \hline
\makecell[l]{Find matching preprocessor conditional, \\
skipping nested ones.} 	                    &   Ctrl+K                  \\ \hline
Select to matching preprocessor conditional.&   Ctrl+Shift+K            \\ \hline
\makecell[l]{Find matching preprocessor conditional backwards, \\
skipping nested ones.} 	                    &   Ctrl+J                  \\ \hline
Select to matching preprocessor conditional backwards. 	&   Ctrl+Shift+J\\ \hline
Previous paragraph. Shift extends selection.&   Ctrl+[                  \\ \hline
Next paragraph. Shift extends selection. 	&   Ctrl+]                  \\ \hline
Previous word. Shift extends selection. 	&   Ctrl+Left               \\ \hline
Next word. Shift extends selection. 	    &   Ctrl+Right              \\ \hline
Previous word part. Shift extends selection.&   Ctrl+/                  \\ \hline
Next word part. Shift extends selection. 	&   Ctrl+\osp               \\ \hline
\caption{Other editor shortcuts}
\end{longtable}
}

\subsection{Files}

{\small 
\begin{longtable}{|l|l|}\hline
\textbf{Function}		        &   \textbf{Shortcut Key}       \\ \hline
\endhead   % To repeat the title line, if needed
New file or project 	        &	Ctrl+N                      \\ \hline
Open existing file or project   &	Ctrl+O                      \\ \hline
Save current file 	            &	Ctrl+S                      \\ \hline
Save all files 		            &	Ctrl+Shift+S                \\ \hline
Close current file 	            &	Ctrl+F4/Ctrl+W              \\ \hline
Close all files 	            &	Ctrl+Shift+F4/Ctrl+Shift+W  \\ \hline
\caption{File shortcuts}
\end{longtable}
}

This is a list of shortcuts provided by the \codeblocks editor component. These shortcuts cannot be rebound.

{\small 
\begin{longtable}{|l|l|}\hline
\textbf{Function}		    &   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Activate next open file 	&   Ctrl+Tab                \\ \hline
Activate previous open file &   Ctrl+Shift+Tab          \\ \hline
\caption{Other file shortcuts}
\end{longtable}
}

\subsection{View}

{\small 
\begin{longtable}{|l|l|}\hline
\textbf{Function}		            &   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Show / hide Messages pane	        &	F2                      \\ \hline
Show / hide Management pane 	    &	Shift+F2                \\ \hline
Move project up (in Project tree) 	&   Ctrl+Shift+Up           \\ \hline
Move project down (in Project tree) &   Ctrl+Shift+Down         \\ \hline
Activate prior (in Project tree)    & 	Alt+F5                  \\ \hline
Activate next (in Project tree)     &	Alt+F6                  \\ \hline
Zoom in / out 	                    &   Ctrl+Roll Mouse Wheel   \\ \hline
Focus editor 	                    &   CTRL+Alt+E              \\ \hline
\caption{View shortcuts}
\end{longtable}
}

\subsection{Search}

{\small 
\begin{longtable}{|l|l|}\hline
\textbf{Function}		    &   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Find 		                &	Ctrl+F                  \\ \hline
Find next 	                &	F3                      \\ \hline
Find previous 	            &	Shift+F3                \\ \hline
Find in files 	            &	Crtl+Shift+F            \\ \hline
Replace 	                &	Ctrl+R                  \\ \hline
Replace in files            &	Ctrl+Shift+R            \\ \hline
Goto line 	                &	Ctrl+G                  \\ \hline
Goto next changed line 	    &	Ctrl+F3                 \\ \hline
Goto previous changed line 	&	Ctrl+Shift+F3           \\ \hline
Goto file 	                &	Alt+G                   \\ \hline
Goto function 	            &	Ctrl+Alt+G              \\ \hline
Goto previous function      &   Ctrl+PgUp               \\ \hline
Goto next function          &   Ctrl+PgDn               \\ \hline
Goto declaration            &   Ctrl+Shift+.            \\ \hline
Goto implementation         &   Ctrl+.                  \\ \hline
Open include file           &   Ctrl+Alt+.              \\ \hline
\caption{Search shortcuts}
\end{longtable}
}

\subsection{Build}

{\small 
\begin{longtable}{|l|l|}\hline
\textbf{Function}		&   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Build 		            &	Ctrl+F9                 \\ \hline
Compile current file	&	Ctrl+Shift+F9           \\ \hline
Run		                &	Ctrl+F10                \\ \hline
Build and Run 	        &	F9                      \\ \hline
Rebuild 	            &	Ctrl+F11                \\ \hline
\caption{Build shortcuts}
\end{longtable}
}

\subsection{Debug}

{\small 
\begin{longtable}{|l|l|}\hline
\textbf{Function}		    &   \textbf{Shortcut Key}   \\ \hline
\endhead   % To repeat the title line, if needed
Debug 	                    &   F8                      \\ \hline
Continue debugging 	        &   Ctrl+F7                 \\ \hline
Step over a code block 	    &   F7                      \\ \hline
Step into a code block 	    &   Shift+F7                \\ \hline
Step out of a code block 	&   Ctrl+Shift+F7           \\ \hline
Toggle breakpoint 	        &   F5                      \\ \hline
Run to cursor 	            &   F4                      \\ \hline
Previous error 	            &   Alt+F1                  \\ \hline
\caption{Debug shortcuts}
\end{longtable}
}