\section{HexEditor}\label{sec:hexeditor}

How a file can be opened in HexEditor within \codeblocks.

\begin{enumerate}
\item \menu{File, Open with HexEditor}
\item Project Navigator context menu (\menu{Open with, Hex editor}
\item Select the Tab Files in the Management Panel. By selecting a file in the FileManager and executing the context menu \menu{Open With Hex editor} this file is opened in HexEditor.
\end{enumerate}

Division of windows:

left is HexEditor view and right is the display as strings

\textbf{Upper line:}
Current position (value in decimal/hex) and percentage (ratio of current cursor position to whole file).

\textbf{Buttons:}

\textbf{Search functions}

\textit{Goto Button:}
Jump to an absolute position. Format in decimal or hex. Relative jump forward or backward by specifying a sign.

\textit{Search:}
Search for hex patterns in the HexEditor view or for strings in the file preview.

\textit{Configuration of the number of columns:}
Exactly, Multiple of, Power of

\textit{Display mode:}
Hex, Binary

\textit{Bytes:}
Select how many bytes should be displayed per column.

\textit{Choice of Endianess:}
BE: Big Endian
LE: Little Endian

\textit{Value Preview:}
Adds an additional view in HexEditor. For a selected value in HexEditor, the value is also displayed as Word, Dword, Float, Double.

\textit{Expression Input:}
Allows you to perform an arithmetic operation on a value in HexEditor. Result of the operation is displayed at the right margin.

\textit{Calc:}
Expression Tester

\textit{Editing a file in the HexEditor:}

Includes Undo and Redo History.

For example, move the cursor to the string view:
Insert spaces with the Insert key.
Delete characters by pressing the Del key.

By entering a text, the existing content is overwritten as a string.

By entering numbers in the HexEditor view the values are overwritten and the preview is updated.

